Asimilación de distintas observaciones

### 2020-12-16 analysis/chapters/experimento_20181122.Rmd
Figuras de RCRV para convencionales

---
title: "Asimilación de distintas observaciones"
output: 
  bookdown::pdf_document2: default
  bookdown::html_document2: default
bibliography: mesoda_ref.bib
---

@@

@@

```{r obs, eval=FALSE, include=FALSE}
tribble(
  ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
  "ADPSFC", "✔", "✔", "✔", "✔",
  "ADPUPA", "✔", "✔", "✔", "✔",
  "AIRCFT", "✔", "✔", "✔", "✔",
  "AIRCAR", "✔", "✔", "✔", "✔",
  "SFCSHP", "✔", "✔", "✔", "✔",
  "ASCATW", "✔", "✔", "✔", "✔", 
  "AUTSFC", " ", "✔", "✔", "✔",
  "SATWND", " ", " ", "✔", "✔",
  "RAD", " ", " ", " ", "✔",
) %>% 
  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
      caption = "Observation types assimilated in each experiment.") %>% 
  kable_classic_2(full_width = F)
```

@@

@@

conv[type %in% c(181, 187) & exp %in% experiments, 
     .(mean_omb = mean(obs.guess, na.rm = TRUE),
                                                   n = .N), 
     by = .(date, exp, usage.flag)] %>% 
  ggplot(aes(date, mean_omb)) +
  geom_line(aes(color = exp)) +
  geom_point(aes(color = exp, size = n)) +
  scale_size_area(max_size = 2) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                          E6 = "SATWND", E7 = "RAD")) +
  facet_wrap(~usage.flag, labeller = labeller(usage.flag = c("1" =  "assimilated",
                                                             "-1" = "rejected"))) +
  labs(y = "Mean of OmB (g/g)", x = NULL,
       color = NULL, size = "n obs") +
  theme_minimal()

@@

fss %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp)) +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                          E6 = "SATWND", E7 = "RAD")) +
  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm"),
                                      w = c("5" = "50 km", "11" = "100 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  labs(color = NULL, x = NULL) +
  theme_minimal()

>>>>

---
title: "Asimilación de distintas observaciones"
output: 
  bookdown::html_document2: default
  bookdown::pdf_document2: default
bibliography: mesoda_ref.bib
---

@@

scale_date <- function(ini, end, break_bin = "6 hour") {
  scale_x_datetime(breaks = seq.POSIXt(ymd_hms(ini), ymd_hms(end), by = break_bin),
                   labels = function(x) {
    fifelse(hour(x) == 0, format(x, "%H\n%b %d"), format(x, "%H"))
  }) 
}

obs.type <- tribble(
  ~type, ~code,
  120, "ADPUPA",
  130, "AIRCFT-AIREP",
  131, "AIRCFT-AMDAR",
  133, "AIRCAR",
  180, "SFCSHP",
  181, "ADPSFC",
  187, "ADPSFC\nno pressure",
  220, "ADPUPA",
  221, "ADPUPA-PIBAL",
  230, "AIRCFT-AIREP",
  231, "AIRCFT-AMDAR",
  233, "AIRCAR",
  280, "SFCSHP",
  281, "ADPSFC",
  287, "ADPSFC\nno pressure",
  290, "ASCATW"
) %>% 
  setDT()

@@

```{r obs}
tribble(
  ~obs_type, ~conv, ~aut, ~satwnd, ~rad,
  "ADPSFC", "x", "x", "x", "x",
  "ADPUPA", "x", "x", "x", "x",
  "AIRCFT", "x", "x", "x", "x",
  "AIRCAR", "x", "x", "x", "x",
  "SFCSHP", "x", "x", "x", "x",
  "ASCATW", "x", "x", "x", "x", 
  "AUTSFC", " ", "x", "x", "x",
  "SATWND", " ", " ", "x", "x",
  "RAD", " ", " ", " ", "x",
) %>% 
  kbl(col.names = c("Obs type", "CONV", "AUT", "SATWND", "RAD"),
      caption = "Observation types assimilated in each experiment.") %>% 
  kable_classic_2(full_width = F)
```

@@

### Ensemble characteristics

#### Reduced Centered Random Variable

```{r rcrv-sfc, fig.cap="Standar deviation of rcrv calculate for surface observations included in boxes of 2.5 degree", fig.scap=c("Wind", "Temperature")}

fread(here("analysis", "data", "derived_data", "uvt_rcrv_box.csv")) %>%   
  .[var %in% c("u", "v") & type %in% c(280, 281, 287, 290)] %>% 
  .[, type.obs := type] %>% 
  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
  geom_raster(aes(fill = sd.y)) +
  geom_mapa() +
  scale_fill_viridis_c(breaks = seq(0, 4, 0.5),
                       limits = c(0, 4),
                       direction = -1,
                       guide = guide_colorstrip(inside = TRUE,
                                                barwidth = 20,
                                                barheight = 0.8)) +
  facet_grid(var~type.obs, 
             labeller = labeller(type.obs = c("280" = "SFCSHP",
                                              "281" = "ADPSFC",
                                              "287" = "ADPSFC\n no pressure",
                                              "290" = "ASCATW"),
                                 var = c("u" = "U wind",
                                         "v" = "V wind" ))) +
  labs(x = "",
       y = "",
       fill = "") +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))

fread(here("analysis", "data", "derived_data", "uvt_rcrv_box.csv")) %>%
  .[var %in% c("t") & type %in% c(180, 181, 187)] %>% 
  .[, type.obs := type] %>% 
  ggplot(aes(ConvertLongitude(lon.box), lat.box)) +
  geom_raster(aes(fill = sd.y)) +
  geom_mapa() +
  scale_fill_viridis_c(breaks = seq(0, 5, 0.5),
                       limits = c(0, 5),
                       direction = -1,
                       guide = guide_colorstrip(inside = TRUE,
                                                barwidth = 20,
                                                barheight = 0.8)) +
  facet_grid(var~type.obs, labeller = labeller(type.obs = c("180" = "SFCSHP",
                                              "181" = "ADPSFC",
                                              "187" = "ADPSFC\n no pressure"),
                                              var = c("t" = "Temperature"))) +
  labs(x = "",
       y = "",
       fill = "") +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.ontop = TRUE,
        panel.grid = element_line(linetype = 3))
```

```{r rcrv-profile, fig.cap="Profiles of mean and standar deviation of rcrv for sounding and aircraft observations."}
obs.type[fread(here("analysis", "data", "derived_data", "rcrv_uvt_perfil.csv")), on = "type"] %>% 
  .[type != 130] %>%
  melt(measure.vars = c("mean.y", "sd.y")) %>% 
  ggplot(aes(nivel.error, value)) +
  geom_hline(yintercept = c(0, 1), color = "grey20") +
  geom_line(aes(color = factor(code), linetype = variable)) +
  scale_color_brewer(palette = "Set1") +
  scale_linetype_manual(values = c("dashed", "solid"),
                        labels = c("Mean RCRV", "SD RCRV")) +
  scale_x_level() +
  coord_flip() +
  facet_wrap(~ var, labeller = labeller(var = c("t" = "Temperature",
                                                "u" = "U wind",
                                                "v" = "V wind"))) +
  labs(y = NULL,
       linetype = "",
       color = NULL) +
 theme_minimal() 
```

@@

conv[type %in% c(181, 187) & exp %in% c("E5", "E7") & usage.flag == 1, 
     .(mean_omb = mean(obs.guess, na.rm = TRUE),
                                                   n = .N), 
     by = .(date, exp, usage.flag)] %>% 
  ggplot(aes(date, mean_omb)) +
  geom_line(aes(color = exp)) +
  geom_point(aes(color = exp, size = n)) +
  scale_size_area(max_size = 2) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                          E6 = "SATWND", E7 = "RAD")) +
  # facet_wrap(~usage.flag, labeller = labeller(usage.flag = c("1" =  "assimilated",
  #                                                            "-1" = "rejected"))) +
scale_date(20181120180000, 20181123120000) +
  labs(y = "Mean of OmB (g/g)", x = NULL,
       color = NULL, size = "number\n of obs") +
  theme_minimal()

@@

fss %>% 
  ggplot(aes(date, fss)) +
  geom_hline(yintercept = 1, color = "darkgray") +
  geom_line(aes(color = exp)) +
  # geom_point(aes(color = exp)) +
  scale_color_manual(values = colores_exp, labels = c(E4 = "CONV", E5 = "AUT", 
                                          E6 = "SATWND", E7 = "RAD")) +
  scale_date(20181121000000, 20181123000000, "12 hours") +
  facet_grid(w~q, labeller = labeller(q = c("1" = "1 mm", "5" = "5 mm", 
                                            "10" = "10 mm", "25" = "25 mm"),
                                      w = c("5" = "50 km", "11" = "100 km", 
                                            "51" = "500 km", "101" = "1000 km"))) +
  labs(color = NULL, x = NULL, y = "FSS") +
  theme_minimal()
